---
title: "The FIFA 2018 World Cup at A Glance"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cerulean
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(plotly)
library(ggplot2)
library(lubridate)
library(tidyr)
library(readr)

#colours to be used
pal = c("#780116ff","#083d77ff","#d8572aff","#3e92ccff","#c32f27ff")
pal2 = c("#083d77ff","#780116ff")
background <- "#d1d1d1ff"
background2 <- "#dbdbdbff"

#Preprocessing of world cup teams data set
worldcup_teams <- read_csv("~/Desktop/RMIT SEM 2 2020/Data Visualisation/Assignment 3/2018_worldcup_v3.csv")
worldcup_teams <- worldcup_teams %>% separate(Datetime, into = c("Date", "Time"), sep = "-")
worldcup_teams$Date <- worldcup_teams$Date %>% parse_date_time(orders = "dmy")
#Selecting required variables into new dataframe
WC_teams <- worldcup_teams %>% select(Date, Stage, `Home Team Name`, `Home Team Goals`, `Away Team Name`, 
                                      `Away Team Goals`)

#group by team to find the most goals by team
WC_grp_goals <- WC_teams %>% group_by(`Home Team Name`, `Away Team Name`)
WC_grp_goals <- WC_grp_goals %>% summarise("total Home goals" = sum(`Home Team Goals`),
                                           "total Away goals" = sum(`Away Team Goals`))

#Filtering out home goals to group by team before adding back with away goals to get total goals scored by team
WC_Home_goals <- WC_grp_goals %>% select(`Home Team Name`, `total Home goals`)
WC_Home_goals <- WC_Home_goals %>% group_by(`Home Team Name`) %>% summarise(totalGoals = sum(`total Home goals`))
WC_away_goals <- WC_grp_goals %>% select(`Away Team Name`, `total Away goals`)
WC_away_goals$`Home Team Name` <- NULL
WC_away_goals <- WC_away_goals %>% group_by(`Away Team Name`) %>% summarise(totalGoals = sum(`total Away goals`))

#renaming team variable to make key to join by
names(WC_Home_goals)[names(WC_Home_goals)=="Home Team Name"] <- "TeamName"
names(WC_away_goals)[names(WC_away_goals)=="Away Team Name"] <- "TeamName"

CountryGoals <- WC_Home_goals %>% left_join(WC_away_goals, by = "TeamName")
CountryGoals$TotalGoals <- CountryGoals$totalGoals.x+CountryGoals$totalGoals.y
CountryGoals$totalGoals.x <- NULL
CountryGoals$totalGoals.y <- NULL

#creating goals Conceded data set
home_goals_Conceded <- WC_grp_goals %>%  select(`Home Team Name`, `total Away goals`)
home_goals_Conceded <- home_goals_Conceded %>%  group_by(`Home Team Name`) %>% 
  summarise(GoalsConceded = sum(`total Away goals`))

away_goals_Conceded <- WC_grp_goals %>% select(`Away Team Name`, `total Home goals`)
away_goals_Conceded <- away_goals_Conceded %>% group_by(`Away Team Name`) %>%
  summarise(goalsConceded = sum(`total Home goals`))

names(home_goals_Conceded)[names(home_goals_Conceded)=="Home Team Name"] <- "TeamName"
names(away_goals_Conceded)[names(away_goals_Conceded)=="Away Team Name"] <- "TeamName"

CountryGoalsConceded <-home_goals_Conceded  %>% left_join(away_goals_Conceded, by = "TeamName")
CountryGoalsConceded$TotalConceded <- CountryGoalsConceded$GoalsConceded + CountryGoalsConceded$goalsConceded
CountryGoalsConceded$GoalsConceded <- NULL
CountryGoalsConceded$goalsConceded <- NULL

#combining goals scored and goals Conceded 
Scored_Conceded <- CountryGoals %>% left_join(CountryGoalsConceded, by = "TeamName")

#creating goals scored to goals Conceded ratio
Scored_Conceded$goalRatio <- round(Scored_Conceded$TotalGoals/Scored_Conceded$TotalConceded,2)

#creating region variable for each country
Asia <- c("IR Iran", 'Japan', "Korea Republic", "Russia", "Saudi Arabia") %>% as.list()
Oceania <- c("Australia") %>% as.list()
Europe <- c("Belgium","Croatia","Denmark","England","France","Germany","Iceland", "Portugal","Serbia","Spain",
            "Sweden", "Switzerland", "Poland") %>% as.list()
Americas <- c("Argentina","Brazil","Colombia", "Costa Rica", "Mexico","Panama", "Uruguay","Peru") %>% as.list()
Africa <- c("Egypt","Morocco","Nigeria","Senegal","Tunisia") %>% as.list()
Scored_Conceded$Region <- NA
Scored_Conceded$Region <- ifelse(Scored_Conceded$TeamName %in% Asia, "Asia", Scored_Conceded$Region)
Scored_Conceded$Region <- ifelse(Scored_Conceded$TeamName %in% Africa, "Africa", Scored_Conceded$Region)
Scored_Conceded$Region <- ifelse(Scored_Conceded$TeamName %in% Americas, "Americas", Scored_Conceded$Region)
Scored_Conceded$Region <- ifelse(Scored_Conceded$TeamName %in% Europe, "Europe", Scored_Conceded$Region)
Scored_Conceded$Region <- ifelse(Scored_Conceded$TeamName %in% Oceania, "Oceania", Scored_Conceded$Region)
#making goals Conceded negative to be able to plot alongside goals scored and then gathering data
Scored_Conceded2 <- Scored_Conceded
Scored_Conceded2$TotalConceded <- Scored_Conceded2$TotalConceded*-1
Scored_Conceded2 <- Scored_Conceded2 %>% gather(TotalGoals, TotalConceded, key = "GoalType", value = "Goals")

#LOADING PLAYERS DATASET 
PlayerGoals <- read_csv("~/Desktop/RMIT SEM 2 2020/Data Visualisation/Assignment 3/PlayerGoals.csv")
PlayerGoals <- PlayerGoals %>% select(Home,Away,Scorer,Type,FinalResult, ScoringTeam)

#LOADING TEAM PERFORMANCE DATASET
teamStats <- read_csv("~/Desktop/RMIT SEM 2 2020/Data Visualisation/Assignment 3/world_cup_2018_stats.csv")
teamStats <- teamStats %>% select(Team, Attempts, `On-Target`, `Distance Covered km`, Tackles, `Yellow cards`,
                                  `Red Cards`, `Fouls Committed`,`Second Yellow Card leading to Red Card`)
```


Goals Breakdown
===================================== 

Column {data-width=750}
-----------------------------------------------------------------------

### Goals Scored By Country {.no-padding}

```{r}
p1 <- ggplot(data = Scored_Conceded2, aes(x = reorder(TeamName,Goals), y = Goals, fill = GoalType))
p1 <- p1 + geom_bar(stat = "identity")+coord_flip()
p1 <- p1 + scale_y_continuous(breaks = c(-10,10), labels = c("10","10"))
p1 <- p1 + scale_fill_manual(values = c("#780116ff","#083d77ff"),labels = c("Goals Conceded","Goals Scored"))
p1 <- p1 + labs(x = "Country", y = "Number of goals", fill = "Goal Type", caption = "Based on data from Kaggle.com (Ghari M 2020)")
p1 <- p1 + theme( panel.background = element_rect(fill = background2), axis.text.x = element_text(), plot.caption = element_text(size = 7, face = "italic"))
p1

```


Column {data-width=850}
-----------------------------------------------------------------------

### Goals Scored and Conceded By Region

```{r}
#preprocessing performed in initial set up chunk
p3 <- plot_ly(data = Scored_Conceded, x = ~TotalGoals, y = ~TotalConceded, color = ~factor(Region),
              colors = pal , size = ~goalRatio,
              text = ~paste("Team: ", TeamName, "<br>Goal Ratio: ", goalRatio))
p3 <- p3 %>% layout(xaxis = list(title = "Goals Scored"), yaxis = list(title = "Goals conceded"))
p3
```

Column {data-width=750}
-----------------------------------------------------------------------

### How Regions Scored Their Goals

```{r}
Scored_Conceded$TeamName <- ifelse(Scored_Conceded$TeamName == "IR Iran", "Iran",Scored_Conceded$TeamName)
Scored_Conceded$TeamName <- ifelse(Scored_Conceded$TeamName == "Korea Republic", "South Korea",Scored_Conceded$TeamName) 
names(PlayerGoals)[names(PlayerGoals)=="ScoringTeam"] <- "TeamName"
Team_Scoring <- Scored_Conceded %>% right_join(PlayerGoals, by = 'TeamName') 
Team_Scoring <- Team_Scoring %>% select(Type, Region)
Team_Scoring$Goals <- 1 
Team_Scoring <- Team_Scoring %>% group_by(Type, Region) %>% summarise(Total = sum(Goals))

p4 <- ggplot(data = Team_Scoring, aes(x = Region, y = Total, fill = Type))
p4 <- p4 + geom_bar(stat = "identity", position = "fill")
p4 <- p4 + scale_fill_manual(values = pal, labels = c("FreeKick", "Inside Penaly Box","Outside Penalty Box", "Own Goal","Penalty"))+labs(y = "Proportion of Goals Scored", caption = "Based on data from Kaggle.com (Šikić F 2020)")
p4 <- p4 + theme( panel.background = element_rect(fill = background2), axis.text.x = element_text(), plot.caption = element_text(size = 7, face = "italic"), panel.grid = element_blank())
p4
```

### Golden Boot (Most Goals Scored by Single Player)

```{r}
names(WC_teams)[names(WC_teams)=="Home Team Name"] <- "Home"
names(WC_teams)[names(WC_teams)=="Away Team Name"] <- "Away"
PlayerGoals2 <- PlayerGoals %>% right_join(WC_teams, by = c("Home", "Away"))
PlayerGoals2$`Home Team Goals` <- NULL
PlayerGoals2$`Away Team Goals`<- NULL
PlayerGoals2$Goals <- 1
PlayerGoals2 <- PlayerGoals2 %>% group_by(Scorer)
PlayerGoals2 <- PlayerGoals2 %>% summarise(GoalsScored = sum(Goals))
PlayerGoals2 <- PlayerGoals2 %>% filter(!is.na(Scorer))
PlayerGoals2 <- PlayerGoals2[order(PlayerGoals2$GoalsScored, decreasing = TRUE),]
PlayerGoals2 <- PlayerGoals2 %>% head(25)

p2 <- ggplot(data = PlayerGoals2, aes(x = reorder(Scorer, GoalsScored), y = GoalsScored,
             fill = ifelse(Scorer == "Kane","A","B")))
p2 <- p2 + geom_bar(stat = "identity", show.legend = FALSE)+coord_flip()+
  scale_fill_manual(values = c("#780116ff","#083d77ff"))
p2 <- p2 + labs(y = "Total Goals Scored", x = "Player Last Name", caption = "Based on Data from kaggle.com (Šikić F 2020)")
p2 <- p2 + theme(panel.background = element_rect(fill = background2), axis.text.x = element_text(), plot.caption = element_text(size = 7, face = "italic"))
p2
```

Team Performance
===================================== 

Column {data-width=750}
-----------------------------------------------------------------------

### Distance Covered By Country across tournament
```{r}
Distance <- teamStats %>% group_by(Team) %>% summarise('Total Distance (Km)'= sum(`Distance Covered km`))

p6 <- ggplot(data = Distance, aes (x = Team, y = `Total Distance (Km)`, 
             fill = ifelse(Team == "France" | Team == "Croatia","A","B")))
p6 <- p6 + geom_bar(stat = "identity", show.legend = FALSE)+scale_fill_manual(values = c("#780116ff","#083d77ff"))
p6 <- p6 + theme( panel.background = element_rect(fill = background2), plot.caption = element_text(size = 7, face = "italic"), axis.text.x = element_text(angle = 90),axis.title.x = element_blank(),)
p6 <- p6+labs(subtitle = "Finalists are highlighted in red", y = "Total Distance Covered (Km)", caption = "Based on data from GITHUB.com (2020)")
p6
```

### Fouls and Yellow Cards Committed by Country
```{r}
Fouls <- teamStats %>% group_by(Team) %>% summarise(YellowCards = sum(`Yellow cards`),
                                                    Fouls = sum (`Fouls Committed`))
Fouls <- Fouls %>%  gather(YellowCards, Fouls, key = "FoulType", value = "NumberComiited")
Fouls$FoulType <- Fouls$FoulType %>% factor(levels = c("Fouls","YellowCards"), labels = c("Fouls",
                                                                                          "Yellow Cards"))

p7 <- ggplot(data = Fouls, aes(x = Team, y = NumberComiited, fill = FoulType))
p7 <- p7 + geom_bar(stat = "identity", position = "dodge", show.legend = FALSE)+facet_grid(FoulType~.,scales = "free")
p7 <- p7 + scale_fill_manual(values = c("#083d77ff","#F4BC1C"))
p7 <- p7 + labs(y = "Total Number Committed", caption = "Based on data from GITHUB.com (2018)")
p7 <- p7 + theme( panel.background = element_rect(fill = background2), plot.caption = element_text(size = 7, face = "italic"), axis.text.x = element_text(angle = 90),axis.title.x = element_blank())
p7
```

Column {data-width=750}
-----------------------------------------------------------------------

### Shots taken versus shots on target
```{r message=FALSE, warning=FALSE}
shotAccuracy <- teamStats %>% group_by(Team) %>% summarise(Total.Taken = sum(Attempts),
                                                           On.Target = sum(`On-Target`))
shotAccuracy <- shotAccuracy %>% mutate(Accuracy = (On.Target/Total.Taken)*100)

p8 <- plot_ly(data = shotAccuracy, x = ~Total.Taken, y = ~On.Target, color = ~Accuracy,
              text = ~paste("Team: ", Team, "<br>Accuracy: ", Accuracy),
              colors = pal2,
              marker = list(size = 10,
                            opacity = 0.8))
p8 <- p8 %>% layout(xaxis = list(title = "Total Shots"), yaxis = list(title = "Shots on target"))
p8
```


References
===================================== 
**References**
 
 * Ghari M 2020, "FIFA World Cup 2018: 2018 FIFA World Cup that was held in Russia, based on FIFA official website.", Retrieved on 16/10/2020, https://www.kaggle.com/rezaghari/fifa-worldcup-2018
 
 * Šikić F 2020, FIFA World Cup 2018 Goals: Every goal and goalscorer from FIFA WC 2018 with additional information, Retrieved on 16/10/2020, https://www.kaggle.com/frankosikic/fifa-world-cup-2018-goals
 
 * DH 2018, "2018-world-cup-stats", Retrieved 21/10/2020 https://gitlab.com/djh_or/2018-world-cup-stats/-/blob/master/world_cup_2018_stats.csv